---
title: "Brian_Singh_DATA607_Wk3"
output: html_document
---

# DATA 607 -- Week Three - R Character Manipulation and Date Processing

### Load libraries and import data
```{r --import-libraries-data}
library(tidyverse)
library(readr)
library(curl)
library(ggplot2)
library(dplyr)

majors<-read.csv(curl("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"))
```

### **1. Identify the majors that contain either "DATA" or "STATISTICS"**
```{r identify-data-statistics-majors}
tibble(Major = majors$Major) %>%
    tidyr::extract(Major,c("Data/Stats"),"(DATA|STATISTICS)",remove=FALSE) %>%
    drop_na()
```
There are two majors containing "Statistics": Management Information Systems and Statistics and Statistics and Decision Science; and one major containing "Data": Computer Programming and Data Processing.

### **2. Transform strings to vector**
```{r transform-strings-vector}
original_strings <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'
original_strings

convert_to_vector <- unlist(str_extract_all(original_strings, pattern = "\"([a-z]+.[a-z]+)\""))
convert_to_vector <- str_remove_all(convert_to_vector, "\"")
convert_to_vector <- convert_to_vector %>%
    str_c(collapse = ", ")
convert_to_vector
```

### **3. Describe, in words, what these expressions will match:**

#### (.)\1\1
This would match any character followed by two repetitions of itself "".

#### "(.)(.)\\2\\1"
This will match two characters, e.g. "xy" as the .., followed by the second character, "y", and then the first character, "x". In this example, xyyx.

#### (..)\1
This would match any two characters repeated once if this was enclosed by "".

#### "(.).\\1.\\1"
This will return a match on a string that has the first character, third and fifth as the same. The 2nd and 4th characters can be any character.

#### "(.)(.)(.).*\\3\\2\\1"
This will match three of the same characters in a row and then any amount of characters, then the first three characters in reverse order.

### **4. Construct regular expressions to match words that:**

#### A: Start and end with the same character.
"^(.)((.*\\1$)|\\1?$)"

#### B: Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
"([A-Za-z][A-Za-z]).*\\1")

#### C. Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
"([A-Za-z][A-Za-z]).*\\1"


