---
title: "Brian_Singh_DATA607_Project1"
output: html_document
---

# DATA 607 Project 1

```{r libraries-import-data}
library(tidyverse)
library(readr)
library(curl)
library(ggplot2)
library(dplyr)
tourney_data<-read.csv(curl("https://raw.githubusercontent.com/brsingh7/DATA607/main/Week4/tournamentinfo.txt"), header=FALSE, sep="|",skip=4)
#tourney_data <- read.csv("tournamentinfo.txt", header=FALSE, sep="|",skip=4)
```

## 2. Separate required data into vectors for use and combine into a data frame with Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating and each round's opponent. Create a separate table with each player's pre-match rating in order to perform a join.
```{r create-data-frames}
rows1 <- tourney_data[seq(1, nrow(tourney_data), 3), ]
rows2 <- tourney_data[seq(2, nrow(tourney_data), 3), ]
player_number <- as.integer(str_match(rows1$V1,"\\d+"))
player_name <- str_trim(rows1$V2)
player_points <- as.double(str_trim(rows1$V3))
player_rating <- as.integer(str_match(str_trim(rows2$V2),"\\s\\d{3,4}+"))
state <- str_trim(rows2$V1)
round1 <- as.integer(str_match(str_trim(rows1$V4),"\\d{1,2}"))
round2 <- as.integer(str_match(str_trim(rows1$V5),"\\d{1,2}"))
round3 <- as.integer(str_match(str_trim(rows1$V6),"\\d{1,2}"))
round4 <- as.integer(str_match(str_trim(rows1$V7),"\\d{1,2}"))
round5 <- as.integer(str_match(str_trim(rows1$V8),"\\d{1,2}"))
round6 <- as.integer(str_match(str_trim(rows1$V9),"\\d{1,2}"))
round7 <- as.integer(str_match(str_trim(rows1$V10),"\\d{1,2}"))

player_df <- data.frame(player_number,player_name,state,player_points,player_rating,round1,round2,round3,round4,round5,round6,round7)
player_ratings_df<-data.frame(player_number,player_rating)
```

### 3. Compare the two tables and return each opponent's rating for each player using a left join.
```{r join-tables}
joined <- left_join(player_df,player_ratings_df,by=c("round1"="player_number"))
joined <- joined %>%
  left_join(player_ratings_df,by=c("round2"="player_number"))%>%
  left_join(player_ratings_df,by=c("round3"="player_number"))%>%
  left_join(player_ratings_df,by=c("round4"="player_number"))%>%
  left_join(player_ratings_df,by=c("round5"="player_number"))%>%
  left_join(player_ratings_df,by=c("round6"="player_number"))%>%
  left_join(player_ratings_df,by=c("round7"="player_number"))
colnames(joined) <- c("Player_ID","Player_Name","State","Player_points","Player_Rating","Round1_Opp","Round2_Opp","Round3_Opp","Round4_Opp","Round5_Opp","Round16_Opp","Round7_Opp","Opp1_Rating","Opp2_Rating","Opp3_Rating","Opp4_Rating","Opp5_Rating","Opp6_Rating","Opp7_Rating")
```

### 4. Calculate Average Pre Chess Rating of Opponents.
```{r calculate-opp-avg}
joined$Opponent_Avg<-rowMeans(joined[,c("Opp1_Rating","Opp2_Rating","Opp3_Rating","Opp4_Rating","Opp5_Rating","Opp6_Rating","Opp7_Rating")],na.rm=TRUE)
joined$Opponent_Avg<-round(joined$Opponent_Avg)
```

### 5. Create final data frame with Player’s Name, Player’s State, Total Number of Points, Player’s Pre-Rating, and Average Pre Chess Rating of Opponents and write to .csv
```{r final-table}
chess_tournament <- joined %>%
  select("Player_Name","State","Player_points","Player_Rating","Opponent_Avg")
head(chess_tournament,15)
write.csv(chess_tournament, "chess_tournament.csv",row.names=FALSE)
#write.csv(chess_tournament,file = file.choose(new = T),row.names=FALSE)
```


