---
title: "Brian_Singh_DATA607_Week5"
author: "Brian Singh"
date: '2022-03-01'
output: html_document
---

# Data 607 Week 5 – Tidying and Transforming Vaccination Data

### 1. Import libraries and data
```{r libraries-import-data}
library(tidyverse)
library(readr)
library(curl)
library(ggplot2)
library(dplyr)
library(DescTools)
#vaccine_data <- read.csv("vaccinations.csv",skip=1,na.strings=c("","NA"))
vaccine_data<-read.csv(curl("https://raw.githubusercontent.com/brsingh7/DATA607/main/Week5/vaccinations.csv"),skip=1,na.strings=c("","NA"))
```

### 2. Tidying and Transforming Data for use in analysis
```{r tidy-data}
#rename columns
colnames(vaccine_data) <- c("Age","Not_Vaccinated","Fully_Vaccinated","Severe_NotVax_Per100K","Severe_FullyVax_Per100K","Efficacy_Vs_Severe_Disease")

#fill in age column "NA"s with value above for categorization
vaccine_data <- vaccine_data %>% 
  fill(Age)

#Extract Age column and values where string ends in % to a new table. Remove irrelevant columns, rename columns
percentages <- vaccine_data[c(2,4),]
percentages <- percentages[,-c(4:6)]
colnames(percentages) <- c("Age","Pct_NotVax","Pct_FullyVax")

#remove rows 2, 4 from vaccine data frame
vaccine_data <- vaccine_data[-c(2,4),]

#join new table with original
vaccine_data2 <- left_join(vaccine_data,percentages,by=c("Age"="Age"))

#convert character values to #s for use in calculations
vaccine_data2$Pct_NotVax <- as.numeric(sub("%","",vaccine_data2$Pct_NotVax))/100
vaccine_data2$Pct_FullyVax <- as.numeric(sub("%","",vaccine_data2$Pct_FullyVax))/100
vaccine_data2$Not_Vaccinated <- as.numeric(gsub(",", "", vaccine_data2$Not_Vaccinated))  
vaccine_data2$Fully_Vaccinated <- as.numeric(gsub(",", "", vaccine_data2$Fully_Vaccinated)) 

#reorder columns
vaccine_data2 <- vaccine_data2[,c("Age","Not_Vaccinated","Pct_NotVax","Severe_NotVax_Per100K","Fully_Vaccinated","Pct_FullyVax","Severe_FullyVax_Per100K","Efficacy_Vs_Severe_Disease")]

vaccine_data2
```

### Q1. Do you have enough information to calculate the total population?  What does this total population represent?

*A: Yes, there is enough information to calculate the total population. The population under 50 years old in the given data accounts for 96.3% of the population under 50, whereas the population of vaxed/nonvaxed over 50 accounts for 98.3% of the population over 50. Given this data, we know there is another 3.7% of the population under 50 and 1.7% of the population over 50, which may represent the population ineligible for the vaccine due to age/health/other factors.*

```{r Q1}
#create a function to calculate populations based on totals and percentages given
full_pop <- function(totalvaxed_unvaxed,total_percent){
    total_pop<-totalvaxed_unvaxed/total_percent
    return(round(total_pop))
}

pop_under_50 <- full_pop(sum(vaccine_data2$Not_Vaccinated[1],vaccine_data2$Fully_Vaccinated[1]),sum(vaccine_data2$Pct_NotVax[1],vaccine_data2$Pct_FullyVax[1]))

pop_over_50 <- full_pop(sum(vaccine_data2$Not_Vaccinated[2],vaccine_data2$Fully_Vaccinated[2]),sum(vaccine_data2$Pct_NotVax[2],vaccine_data2$Pct_FullyVax[2]))

total_pop <- pop_under_50+pop_over_50
print(paste("The total population in Israel is",total_pop,"people."))
```

### Q2. Calculate the Efficacy vs. Disease; Explain your results.

*A: Based on the calculations, the vaccine has a 74% efficacy against contracting severe disease for the population under 50 years old. Interestingly, the vaccine appears to have a negative effect on the population over 50 years old as we see an increase in severe cases for those fully vaccinated by approximately 0.12% (290 fully vaxed vs. 171 unvaxed). In totality, regardless of age, the vaccine appears to have a negative effect, however this is skewed by the large disparity in the effectiveness between the two age groups.*

```{r Q2}
#create a function to calculate efficacy
efficacy_severe <- function(vaxed_severe,unvaxed_severe){
    pct_severe<-(vaxed_severe/100000)/(unvaxed_severe/100000)
    return(1-pct_severe)
}

efficacy_under50 <- round(efficacy_severe(vaccine_data2$Severe_FullyVax_Per100K[1],vaccine_data2$Severe_NotVax_Per100K[1]),3)
print(paste("The efficacy under 50 years old is",efficacy_under50*100,"%"))


efficacy_over50 <- round(efficacy_severe(vaccine_data2$Severe_FullyVax_Per100K[2],vaccine_data2$Severe_NotVax_Per100K[2]),3)
print(paste("The efficacy over 50 years old is",efficacy_over50*100,"%"))

total_efficacy <- round(efficacy_severe((vaccine_data2$Severe_FullyVax_Per100K[1]+vaccine_data2$Severe_FullyVax_Per100K[2]),(vaccine_data2$Severe_NotVax_Per100K[1]+vaccine_data2$Severe_NotVax_Per100K[2])),3)
print(paste("The efficacy under 50 years old is",total_efficacy*100,"%"))
```

### Q3. From your calculation of efficacy vs. disease, are you able to compare the rate of severe cases in unvaccinated individuals to that in vaccinated individuals?

*Yes. For individuals under 50, the rate of severe cases decreases by 0.032% for those vaccinated, however, it increases by 0.12% for those over 50. This is the basis of the results of the efficacy calculation.*
```{r Q3}
#calculate rate of change between severe cases for vaxed and unvaxed individuals
rate_severe_vaxed_unvaxed <- function(vaxed_severe,unvaxed_severe){
    rate<-(vaxed_severe-unvaxed_severe)/100000*100
    return(rate)
}

rate_under50 <- round(rate_severe_vaxed_unvaxed(vaccine_data2$Severe_FullyVax_Per100K[1],vaccine_data2$Severe_NotVax_Per100K[1]),3)
print(paste("The rate of severe cases in vaccinated vs. unvaccinated individuals under 50 years old is",rate_under50,"%"))

rate_over50 <- round(rate_severe_vaxed_unvaxed(vaccine_data2$Severe_FullyVax_Per100K[2],vaccine_data2$Severe_NotVax_Per100K[2]),3)
print(paste("The rate of severe cases in vaccinated vs. unvaccinated individuals over 50 years old is",rate_over50,"%"))
```